
\section{Introduction} \label{into}

The estimation and tracking of 3D human poses is an important research area in computer vision, driven by its broad range of immediate applications across multiple domains. For example, in robotics and human-robot interactions it can find application in industrial environments and collaborative workspaces where robots must accurately perceive human poses and predict movements to maintain safe operational boundaries and avoid collisions with dangerous machinery. Similarly, for crowd control tracking of multi-person trajectories in public spaces facilitates anomalous behavior detection, crowd density planning, and automated surveillance. Other impactful applications can be found in healthcare, such as gait analysis, automated fall detection, and remote patient monitoring. Finally, human pose estimation and tacking can find applications also in the entertainment domain, such as the use real-time motion capture for virtual avatars or natural body-based control interfaces for gaming.

Despite the maturity of real-time 2D human pose estimation, transferring these methods to a robust and real-time capable 3D setting remains a challenging task due to issues such as depth ambiguities, severe self-occlusions, and the high perspective variance of human motion. The primary objective of this project is to design, implement, and evaluate a computer vision architecture capable of tracking the skeletal movement of multiple people from 2D pose estimations extracted from multiple views by off-the-shelf 2D human pose estimation models. The system should be able to fuse the information gained from the 2D video frames of multiple camera viewpoints into coherent and temporally consistent 3D pose trajectories.

For the evaluation of the result, this project will use a subset of the CMU Panoptic dataset \parencite{Joo_2015_ICCV,Joo_2017_TPAMI}, which contains videos captured from multiple calibrated cameras together with ground truth pose and tracking information for each frame. Using this evaluation set this project further aims to investigate the impact of model capacity for the 2D pose estimation model, as well as the impact of viewpoint count on both the achieved tracking accuracy and the runtime efficiency of the system.

The remainder of this report is structured as follows: \Cref{dataset-desc} describes the CMU Panoptic dataset that has been used for the evaluation of this project, explaining its structure and characteristics. \Cref{background} reviews the relevant theoretical background, referencing the course material where appropriate. \Cref{methodology} details the architecture of the tracking system, as well as the evaluation strategy including evaluation metrics. \Cref{results} presents the quantitative and qualitative evaluation results. Finally, \Cref{conclusion} concludes the report with a summary of the findings and outlines limitations, highlighting potential areas for further improvements.
