
\input{tikzfigures/architecture.tex}

\section{Methodology} \label{methodology}

This section of the report will detail the architectures of the tracking system developed in the project. Further it covers the selected evaluation metrics and how evaluation was performed.


\subsection{System Architecture} \label{arch}

This subsection will go into detail about the internal implementation of the tracking system. \Cref{fig:architecture} show the overall pipeline used in this projects tracking algorithm.


\subsubsection{Overview} \label{arch-overview}

The core of the tracking pipeline runs a recursive estimation loop on incoming multi-view video frames. The input to the system is a number of synchronized multi-view video feeds, and the final result is a 3D representation of the skeletal joint of each person in the scene.

First, multi-view videos are passed through the YOLO detection network, yielding 2D bounding boxes and keypoints. Concurrently, the tracking module predicts the current 3D state of all previously known tracks based on continuous physics models. The keypoints of the predicted tracks are projected into the 2D camera planes and matched against the newly observed 2D detections. Detections matched to existing tracks trigger an extended Kalman filer update for the matched track to refine its 3D state. Unmatched, novel 2D detections matched between views, and if at least two views contain the new target, triangulated to instantiate new 3D tracks. Finally, updated tracks and newly created tracks are merged, and the loop repeats for the next timestamp.


\subsubsection{2D Detections and Uncertainty Estimation} \label{arch-detect}

During inference, images from each active camera view are batched and passed to the YOLO26-pose detector using PyTorch \parencite{paszke2019pytorch,ansel2024pytorch}. The detector outputs 2D keypoint positions along with visibility confidence scores $c_k \in [0, 1]$. As a first filter we select only detections that have a confidence above $\tau_{obj}$. Further, we filter out any detections for which fewer than $N_{kpt}$ are visible with a confidence above $\tau_{kpt}$. However, for performing the Kalman filter updates we need a probabilistic framework in which we model the prediction uncertainty as Gaussian noise. This need can not be satisfied by the confidence score alone. To estimate the measurement noise covariance dynamically from the outputs of the YOLO model, the system relies on bounding box and visibility scaling. The variance $\sigma_k^2$ for each detected keypoint $k$ is computed dynamically based on the bounding box magnitude and the joint visibility confidence. The specific formulation used is:
\begin{align*}
    \sigma_k^2 &= \sigma_{min}^2 + d_{bb}^2 \cdot \left( \frac{\sigma_{vis}^2}{\max(0, \bar{c}_k) + \frac{\sigma_{vis}^2}{\sigma_{inv}^2}} \right)
\end{align*}
where $\bar{c}_k = \frac{c_k - \tau_{kpt}}{1.0 - \tau_{kpt}}$ is the confidence score normalized above the tuned keypoint threshold $\tau_{kpt}$, $d_{bb}^2$ is the length of the diagonal of the bounding box, and $\sigma_{min}, \sigma_{vis}, \sigma_{inv}$ are tuning parameters specifying respectively the minimum variance, minimum bounding box dependent variance, and maximum bounding box dependent variance. This ensures that occluded joints are appropriately penalized with high uncertainty while visible joint get a lower variance. The keypoint detections are assumed to all be independent, so the final measurement noise covariance matrix is build by filling the diagonals.

Further, before passing the resulting 2D keypoints to the rest of the tracking system, the 2D pixel coordinates are transformed into normalized camera coordinates utilizing the camera intrinsics and distortion coefficients as follows:
\begin{align*}
    \mathbf{x}_{pixel} &= \begin{bmatrix} u \\ v \end{bmatrix} \xrightarrow{\mathbf{K}^{-1}} \mathbf{x}_{distorted} \xrightarrow{\text{Iterative Undistortion}} \mathbf{x}_{normalized}
\end{align*}
To maintain consistent uncertainty representation, we also transform the observation covariance matrix $\Sigma_{pixel}$ for the 17 keypoints into normalized space using the following approximation:
\begin{align*}
    \Sigma_{norm} &\approx K^{-1} \Sigma_{pixel} K^{-T}
\end{align*}
The main motivation behind this approach is that it simplifies the triangulation for new tracks as well as the observation process modeling in the Kalman filter update for old tracks by removing the non-linear distortion effects. As an additional benefit, it becomes more efficient to compute not only the Kalman filter updates but also the computation of cost matrices for the association step, which need to perform projection from the internal 3D state.


\subsubsection{State Representation and Kalman Filtering} \label{arch-filter}

In the Kalman filter implementation of the project, the state for each person is tracked separately. The state vector is composed by modeling the human pose as a collection of 17 keypoints. The state vector $\mathbf{x} \in \mathbb{R}^{102}$ encapsulates the 3D coordinates and the 3D velocities for all 17 keypoints resulting in a state vector shaped as:
\begin{align*}
    \mathbf{x} &= \begin{bmatrix} \mathbf{p}_1^T, \; \dots, \; \mathbf{p}_{17}^T, \; \mathbf{v}_1^T, \; \dots, \; \mathbf{v}_{17}^T \end{bmatrix}^T
\end{align*}
where $\mathbf{p}_k = [x_k, y_k, z_k]^T \in \mathbb{R}^3$ and $\mathbf{v}_k = [\dot{x}_k, \dot{y}_k, \dot{z}_k]^T \in \mathbb{R}^3$ are the 3D position and velocity vectors for keypoint $k$.

The state dynamics are modeled by a continuous-time linear velocity process. The decision was made to specify the dynamics in the continuous domain in order to be able to handle variable frame rates which could for example arise due to frames being dropped during computationally more demanding scenes. The continuous dynamics matrix $\mathbf{F}$ and process noise $\mathbf{Q}$ are mapped to discrete time steps $\Delta t_i$ dynamically for each frame $i$ using a second-order Taylor expansion based approximation:
\begin{align*}
    \mathbf{F}_k &\approx \mathbf{I} + \mathbf{F} \Delta t_i + \frac{1}{2} \mathbf{F}^2 \Delta t_i^2 \\
    \mathbf{Q}_k &\approx \mathbf{Q} \Delta t_i + \frac{1}{2} (\mathbf{F} \mathbf{Q} + \mathbf{Q} \mathbf{F}^T) \Delta t_i^2
\end{align*}

The measurement function $h_c(\mathbf{x}): \mathbb{R}^{102} \to \mathbb{R}^{34}$ acts as a non-linear pinhole projection mapping the 102-dimensional 3D state back to a 34-dimensional observation of 2D coordinates in camera $c$:
\begin{align*}
    h_c(\mathbf{x}) &= \begin{bmatrix} \pi_c(\mathbf{p}_1)^T & \pi_c(\mathbf{p}_2)^T & \dots & \pi_c(\mathbf{p}_{17})^T \end{bmatrix}^T
\end{align*}
where $\pi_c(\mathbf{p}_k)$ is the perspective pinhole projection of the 3D point into normalized camera coordinates:
\begin{align*}
    \pi_c(\mathbf{p}_k) &= \frac{1}{Z_{c,k}} \begin{bmatrix} X_{c,k} \\ Y_{c,k} \end{bmatrix} = \frac{1}{\mathbf{R}_{c,3} \mathbf{p}_k + t_{c,3}} \begin{bmatrix} \mathbf{R}_{c,1} \mathbf{p}_k + t_{c,1} \\ \mathbf{R}_{c,2} \mathbf{p}_k + t_{c,2} \end{bmatrix}
\end{align*}
Because multiple cameras observe the target simultaneously, the system concatenates these independent multi-camera views into a single centralized extended Kalman filter observation update instead of applying them sequentially. For a track observed by $M$ cameras the observations are merged as follows:
\begin{align*}
    \mathbf{z}_{merged} &= \begin{bmatrix} \mathbf{z}_1^T \vdots \mathbf{z}_M^T \end{bmatrix}^T \in \mathbb{R}^{34M} \\
    \Sigma_{merged} &= \text{diag}(\Sigma_1, \dots, \Sigma_M) \in \mathbb{R}^{34M \times 34M}
\end{align*}
This formulation can handle asynchronous data naturally by simply omitting cameras without active detections.

Because the projection mapping is non-linear, the extended Kalman filter requires the computation of the Jacobian $\mathbf{H}_i \in \mathbb{R}^{34M \times 102}$. Rather than deriving and hard-coding this computation analytically, the system leverages the automatic differentiation capabilities of PyTorch \parencite{paszke2019pytorch,ansel2024pytorch} to compute the exact measurement Jacobian dynamically at runtime.


\subsubsection{Data Association} \label{arch-assoc}

Data association is performed in two separate steps. First, the detections for in view are matched against the existing 3D track predictions. This is performed independently for each view. Next, all detections that were not matched to any of the existing tracks must be matched between views in order to find opportunities for triangulation.

The association between detections and existing tracks is performed independently for each view by first projecting the prediction to the 2D camera plane. Note that we also project the 3D covariances using the first-order approximation using the Jacobian. A cost matrix is populated between projected predictions and YOLO derived detections. The distance metric is a modified Mahalanobis distance that adds the log-determinant of the combined covariance to discourage highly uncertain matches with low statistical power:
\begin{align*}
    d_\Sigma(\mathbf{z}, \hat{\mathbf{z}}) &= (\mathbf{z} - \hat{\mathbf{z}})^T \Sigma^{-1} (\mathbf{z} - \hat{\mathbf{z}}) \\
    C_{i,j} &= d_{\Sigma_{pred} + \Sigma_{obs}}(\mathbf{z}, \hat{\mathbf{z}}) + \log|\Sigma_{pred} + \Sigma_{obs}|  - N_{dim} \theta_{mo}
\end{align*}
where $N_{dim} = 34$, and $\theta_{mo}$ is a threshold parameter. Note that we fill the cost matrix with zero padding to allow for unmatched tracks or detections. The optimal association is found by solving this cost matrix using the Hungarian algorithm  \parencite{kuhn1955hungarian} via the SciPy library \parencite{virtanen2020scipy} implementation. This process can be interpreted as an implementation of maximum likelihood based model selection amongst all possible associations.

Detections that are not matched to existing tracks by the above method may belong to new people entering the scene. Since these detections lack 3D track priors, we must match them across different camera views. To achieve this goal, this project employs an iterative greedy-optimal matching approach to associate 2D detections across views. The algorithm works by first building associations between the first two views, then iteratively adding views. During the iterative addition step we compute a cost matrix and solve for the optimal association using the Hungarian algorithm in an effort to minimize the reprojection error:
\begin{align*}
    C_{cross} &= \frac{1}{K}\sum_{k=1}^K \left( d_{\Sigma_k}(\mathbf{z}_k, \pi_k(\mathbf{X}_{3D})) + \ln|\Sigma_k| \right) - N_{dim}\theta_{mn}
\end{align*}
where $\mathbf{X}_{3D}$ is the triangulated 3D point from the $K$ associated views, and $\theta_{mn}$ is the new-track association threshold. Detections matched in at least two views are grouped to initialize a new track via triangulation.


\subsubsection{Track Creation and Deletion} \label{arch-tracks}

To prevent spurious detections from creating false tracks, we manage the lifecycle of each track using a state machine. New tracks that have just been added via triangulation of detections start out as tentative tracks.  If a tentative track is not matched to at least one detection in the next frame, it is deleted immediately. Once a tentative track have been observed for $\ge N_{age}$ consecutive frames, it is promoted to an active tracks. Active tracks are more robust to occasional missing detections. If an active track is not observed, its state is updated only using the Kalman filter prediction step. However, if it remains unmatched for more than $N_{inv}$ consecutive frames, it is deleted.

\begin{table*}[!ht]
  \centering
  \begin{tabular}{|r|r|c|c|c|c|c|c|c|}
  \hline
    Model & \# Cameras & Detection & Matching & EKF Update & EKF Prediction & Total \\ \hline
    \multirow{4}{*}{YOLO26n-pose}  & 4 & 0.0445 & 0.0166 & 0.0080 & 0.0005 & 0.0697 \\ \cline{2-7}
      & 8 & 0.0840 & 0.0313 & 0.0127 & 0.0006 & 0.1288 \\ \cline{2-7}
      & 12 & 0.1324 & 0.0431 & 0.0165 & 0.0006 & 0.1929 \\ \cline{2-7}
      & 16 & 0.1805 & 0.0567  & 0.0210 & 0.0007 & 0.2592 \\ \hline \hline
    YOLO26s-pose & 4 & 0.0876 & 0.0178 & 0.0083 & 0.0005 & 0.1145 \\ \hline
    YOLO26m-pose & 4 & 0.2519 & 0.0177 & 0.0083 & 0.0005 & 0.2786 \\ \hline
    YOLO26l-pose & 4 & 0.2613 & 0.0177 & 0.0083 & 0.0005 & 0.2879 \\ \hline
  \end{tabular}
  \caption{Average per-frame runtime of the tracking system for different camera and model configurations. Times are given in seconds.}
  \label{fig:perf}
\end{table*}

\subsection{Evaluation Procedure} \label{eval-proc}

For this project the system has been evaluated by comparing its 3D tracking predictions frame-by-frame against the ground truth annotations in the CMU Panoptic dataset. The experiments were run across all five selected sequences mentioned in \cref{dataset-desc}, evaluating various combinations of YOLO pose estimation models, testing \inlinecode{YOLO26n-pose}, \inlinecode{YOLO26s-pose}, \inlinecode{YOLO26m-pose}, and \inlinecode{YOLO26l-pose}, and spacial view densities of 4, 8, 12, and 16. Note that due to computation demands, only the \inlinecode{YOLO26n-pose} model has been tested with more than 4 cameras. All evaluations were performed on AMD Ryzen AI 9 HX 370 CPU with most PyTorch \parencite{paszke2019pytorch,ansel2024pytorch} operation running on the integrated AMD Radeon 890M GPU.

Due to the scope of the project, system parameter of the system were initialized heuristically rather than via exhaustive grid-search or end-to-end learning. While these rigid parameter values are functional, an identified area for improvement is optimizing them directly from tracking data either using grid search or end-to-end learning.

Detection thresholds for the YOLO based detector we set to $\tau_{obj} = 0.5$ and $\tau_{kpt} = 0.25$, using also $N_{kpt} = 3$ as the minimum number of visible keypoints. The minimum pixel variance has been set to $\sigma_{min}^2 = 16.0$ pixels$^2$, with visibility and invisibility scale variances set to $\sigma_{vis}^2 = 0.005$ and $\sigma_{inv}^2 = 5$. The matching thresholds for both cross-view and prediction-view matching have both been set to $\theta_{mn} = \theta_{mo} = 0$. For the tracking lifecycle we configured the age for tentative track promotion to $N_{age} = 3$ and the maximum duration without detections for active tracks to $N_{inv} = 10$. Finally, the continuous physic matrices have been set as follows:
\begin{align*}
    \mathbf{F} = \begin{bmatrix} \mathbf{0}_{51 \times 51} & \mathbf{I}_{51 \times 51} \\ \mathbf{0}_{51 \times 51} & -0.5 \cdot \mathbf{I}_{51 \times 51} \end{bmatrix}, & \;
    \mathbf{Q} = \begin{bmatrix} \sigma_{p}^2 \cdot \mathbf{I}_{51 \times 51} & \mathbf{0}_{51 \times 51} \\ \mathbf{0}_{51 \times 51} & \sigma_{v}^2 \cdot \mathbf{I}_{51 \times 51} \end{bmatrix}
\end{align*}
with $\sigma_{v}^2 = (2 \frac{m}{s})^2$ and $\sigma_{p}^2 = (5 cm)^2$. This matrix models constant velocity dynamics with a damping term of $-0.5$ on velocity to stabilize motion predictions.


\subsection{Evaluation Metrics} \label{eval-metrics}

To measure tracking and reconstruction performance, this project makes use of standard metrics from the Multi-Object Tracking and 3D Pose Estimation literature \parencite{bernardin2008evaluating,ionescu2013human3}. All measures are reported in centimeters, and we use a mean joint error of 15 centimeters as the threshold for matches for all metrics except MPJPE.

\paragraph{Mean Per-Joint Position Error (MPJPE)}
MPJPE measures the average 3D Euclidean distance between predicted and ground truth keypoints across all matched frames and joints.
\begin{align*}
    \text{MPJPE} &= \frac{1}{N_{matched}} \sum_{i=1}^{N_{matched}} \left( \frac{1}{17} \sum_{k=1}^{17} \|\mathbf{p}_{k,i}^{pred} - \mathbf{p}_{k,i}^{gt}\|_2 \right)
\end{align*}
For this metric we do not apply a matching threshold, instead matching all predictions against the best matching ground truth track via Hungarian method based assignment.

\paragraph{Multi-Object Tracking Accuracy (MOTA)}
MOTA evaluates overall tracking performance by accounting for three types of errors, false positives, false negatives, and identity switches:
\begin{align*}
    \text{MOTA} &= 1.0 - \frac{\sum_{t} \text{FN}_t + \text{FP}_t + \text{IDS}_t}{\sum_{t} \text{GT}_t}
\end{align*}
where $\text{FN}_t$ is the number of false negatives, $\text{FP}_t$ the number of false positives, $\text{IDS}_t$ that of identity switches, and $\text{GT}_t$ is the number of ground truth individuals in frame $t$.

\paragraph{Multi-Object Tracking Precision (MOTP)}
MOTP measures the alignment of tracked objects by calculating the average distance error for correctly matched targets. Note that in contrast to MPJPE this metric is only computed over the true positives matched within the 15 centimeters maximum threshold.
\begin{align*}
    \text{MOTP} &= \frac{\sum_{t,d} \frac{1}{17} \sum_{k=1}^{17} \|\mathbf{p}_{t,d,k}^{pred} - \mathbf{p}_{t,d,k}^{gt}\|_2}{\sum_{t} \text{TP}_t}
\end{align*}
where $\text{TP}_t$ is the number of true positives.

\paragraph{Precision and Recall}
Finally, we also report the standard precision and recall metrics.
\begin{align*}
    \text{Precision} = \frac{\sum_t \text{TP}_t}{\sum_t \text{TP}_t + \text{FP}_t}, &\quad \text{Recall} = \frac{\sum_t \text{TP}_t}{\sum_t \text{TP}_t + \text{FN}_t}
\end{align*}
